---
title: "Structural Bioinformatics (pt2)"
author: "Yerin Go A16272901"
format: pdf
editor: visual
---

AlphaFold has changed the game for protein structure prediction and allows anyone with bioinfo skills to predict the structure of virtually any protein 

We ran via GoogleColab at: 
https://github.com/sokrypton/ColabFold 

In particular we used their AlphaFold2_mmseq2 version that uses mmseq2 rather than HMMer for sequence search. 

The main outputs include a set of **PDB structure** along with matching **JSON format files** that tell us how good the resulting models might be. 

Lets start by loading these structures in Mol* 

```{r}
library(bio3d)

# Change this for YOUR results dir name
results_dir <- "test_94b5b" 
```

```{r}
# File names for all PDB models
pdb_files <- list.files(path=results_dir,
                        pattern="*.pdb",
                        full.names = TRUE)

# Print our PDB file names
basename(pdb_files)
```
```{r}
library(bio3d)

# Read all data from Models 
#  and superpose/fit coords
pdbs <- pdbaln(pdb_files, fit=TRUE, exefile="msa")
```
```{r}
pdbs
```

```{r}
rd <- rmsd(pdbs, fit=T)
```

```{r}
range(rd)
```

```{r}
library(pheatmap)
colnames(rd) <- paste0("m",1:5)
rownames(rd) <- paste0("m",1:5)
pheatmap(rd)
```
```{r}
# Read a reference PDB structure
pdb <- read.pdb("1hsg")
```
```{r}
plotb3(pdbs$b[1,], typ="l", lwd=2, sse=pdb)
points(pdbs$b[2,], typ="l", col="red")
points(pdbs$b[3,], typ="l", col="blue")
points(pdbs$b[4,], typ="l", col="darkgreen")
points(pdbs$b[5,], typ="l", col="orange")
abline(v=100, col="gray")
```
```{r}
core <- core.find(pdbs)
```
```{r}
core.inds <- print(core, vol=0.5)
```
```{r}
xyz <- pdbfit(pdbs, core.inds, outpath="corefit_structures")
```
```{r}
rf <- rmsf(xyz)

plotb3(rf, sse=pdb)
abline(v=100, col="gray", ylab="RMSF")
```
If the predicted model has more than one domain, each domain may have higher confidence, yet the relative positions of the domain may not. The estimated reliability of relative domain positions is in graphs of predicted aligned errors (PAE) which are included in the downloaded zip files and analyzed in R above. 




